---
title: "Homework 2"
author: "Jianyou Liu"
date: "September 30, 2018"
output: github_document
---

```{r setup_load_packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
# Problem 1

### The code chunk below imports the NYC Subway Station data

```{r import_transit_data}
raw_transit_data = read_csv(file = "./hw2_data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")


```
### The code chunk below cleans the NYC Subway Station data

```{r clean1_transit_data}
tidy1_transit_data = 
  janitor::clean_names(raw_transit_data) %>%
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada ) 
  

```
### The code chunk below converts the "entry" variable into logical
```{r convert_variable, eval = FALSE}
recode(tidy1_transit_data$entry, YES = TRUE, NO = FALSE)
```
The dataset contains **line**, **station names**, **station latitude/longitude**, **routes served**, **whether or not there is an entry**, **whether or not there is a vending machine**, **entrance type**, and **Whether ADA compliant** as its variables.

So far, I cleaned the data by first converting the variables names from uppercase to lowercase, then selected those variables of interest. The data is still not tidy right now.

The *dimension* of the resulting data set is `r dim(tidy1_transit_data)`

### The code chunk below further cleans the dataset
```{r clean2_transit_data}
tidy2_transit_data = distinct(tidy1_transit_data, line, station_name, .keep_all = TRUE)

ada_data = filter(tidy2_transit_data, ada == TRUE)

proportion_data = filter(tidy2_transit_data, vending == "NO")
```
*   The number of distinct stations is equal to `r nrow(tidy2_transit_data)`.
*   `r nrow(ada_data)`stations are ADA compliant.
*   `r table(proportion_data$entry)["YES"]/nrow(proportion_data)` of station entrances/exits without vending allow entrance.

