---
title: "Homework 2"
author: "Jianyou Liu"
date: "September 30, 2018"
output: github_document
---

```{r setup_load_packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
# Problem 1

### The code chunk below imports the NYC Subway Station data

```{r import_transit_data}
raw_transit_data = read_csv(file = "./hw2_data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")

raw_transit_data
```
### The code chunk below cleans the NYC Subway Station data

```{r clean1_transit_data}
tidy1_transit_data = 
  janitor::clean_names(raw_transit_data) %>%
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada ) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

tidy1_transit_data
```

The dataset contains **line**, **station names**, **station latitude/longitude**, **routes served**, **whether or not there is an entry**, **whether or not there is a vending machine**, **entrance type**, and **Whether ADA compliant** as its variables.

So far, I cleaned the data by first converting the variables names from uppercase to lowercase. Then, I selected those variables of interest. Finally, I converted the *entry* variable from character to logical. The data is still not tidy right now.

The *dimension* of the resulting data set is `r dim(tidy1_transit_data)`

### The code chunk below further cleans the dataset
```{r clean2_transit_data}
tidy2_transit_data = distinct(tidy1_transit_data, line, station_name, .keep_all = TRUE)

ada_data = filter(tidy2_transit_data, ada == TRUE)

proportion_data = filter(tidy2_transit_data, vending == "NO")

tidy2_transit_data
```
*   The number of distinct stations is equal to `r nrow(tidy2_transit_data)`.
*   `r nrow(ada_data)` stations are ADA compliant.
*   `r table(proportion_data$entry)["TRUE"]/nrow(proportion_data)` of station entrances/exits without vending allow entrance.(Using the dinstict stations dataset)

### The following code chunk reformats data to allow "route number" and "route name" be distinct variables

```{r reformat_data}
reformat_transit_data = 
  gather(tidy2_transit_data, key = route_number, value = route_name, route1:route11) %>% 
  select(line, station_name, route_number, route_name, everything())

reformat_transit_data

A_train_data = filter(reformat_transit_data, route_name == "A")

```
The number of distinct stations that serve train A is `r nrow(A_train_data)`.

Of the stations that serve the A train, `r table(A_train_data$ada)["TRUE"]` are ADA compliant.

